References

Achdou, Yves, Jiequn Han, Jean-Michel Lasry, Pierre-Louis Lions, and Benjamin Moll, “Income and Wealth Distribution in Macroeconomics: A Continuous-Time Approach,” The Review of Economic Studies, 04 2021, 89 (1), 45–86.

Ahn, SeHyoun, Greg Kaplan, Benjamin Moll, Thomas Winberry, and Christian Wolf,

“When Inequality Matters for Macro and Macro Matters for Inequality,” NBER Macroeconomics Annual, 2018, 32 (1), 1–75.

26 See, for example, Niv (2009), Glimcher (2011), Caplin and Dean (2008), Glimcher et al. (2013), Gershman and

Daw (2017), and Barberis and Jin (2023).

Aiyagari, S. Rao, “Uninsured Idiosyncratic Risk and Aggregate Saving,” The Quarterly Journal

of Economics, August 1994, 109 (3), 659–84.

Auclert, Adrien, Bence Bardóczy, Matthew Rognlie, and Ludwig Straub, “Using the

Sequence-Space Jacobian to Solve and Estimate Heterogeneous-Agent Models,” Econometrica, September 2021, 89 (5), 2375–2408.

, Matthew Rognlie, and Ludwig Straub, “Fiscal and Monetary Policy with Heterogeneous

Agents,” Annual Review of Economics, 2025, 17 (Volume 17, 2025), 539–562.

, Rodolfo Rigato, Matthew Rognlie, and Ludwig Straub, “Beyond Certainty Equivalence:

Second-order Dynamics in the Sequence Space,” Presentation slides 2025.

Azinovic, Marlon, Luca Gaegauf, and Simon Scheidegger, “Deep Equilibrium Nets,” International Economic Review, 2022, 63 (4), 1471–1525.

Azinovic-Yang, Marlon and Jan Žemlička, “Deep Learning in the Sequence Space,” arXiv

preprint arXiv:2509.13623, 2025.

Barberis, Nicholas and Lawrence Jin, “Model-free and Model-based Learning as Joint Drivers

of Investor Behavior,” NBER Working Papers 31081 March 2023.

Bilal, Adrien, “Solving Heterogeneous Agent Models with the Master Equation,” NBER Working Papers 31103, National Bureau of Economic Research 2023.

Bradbury, James, Roy Frostig, Peter Hawkins, Matthew James Johnson, Chris Leary, Dougal

Maclaurin, George Necula, Adam Paszke, Jake VanderPlas, Skye Wanderman-Milne, and

Qiao Zhang, “JAX: composable transformations of Python+NumPy programs,” 2018.

Branch, William A., “Restricted Perceptions Equilibria and Learning in Macroeconomics,”

Post Walrasian Macroeconomics: Beyond the Dynamic Stochastic General Equilibrium Model. Cambridge University Press, New York, 2006, pp. 135–160.

Bray, Margaret, “Learning, Estimation, and the Stability of Rational Expectations,” Journal of

Economic Theory, April 1982, 26 (2), 318–339.

Calvano, Emilio, Giacomo Calzolari, Vincenzo Denicolo, and Sergio Pastorello, “Artificial

intelligence, algorithmic pricing, and collusion,” American Economic Review, 2020, 110 (10),

3267–3297.

Caplin, Andrew and Mark Dean, “Dopamine, Reward Prediction Error, and Economics,” The

Quarterly Journal of Economics, 2008, 123 (2), 663–701.

Cardaliaguet, Pierre, François Delarue, Jean-Michel Lasry, and Pierre-Louis Lions, The Master Equation and the Convergence Problem in Mean Field Games, Vol. 201 of Annals of Mathematics

Studies, Princeton University Press, 2019.

Chen, Mingli, Andreas Joseph, Michael Kumhof, Xinlei Pan, and Xuan Zhou, “Deep Reinforcement Learning in a Monetary Model,” 2023.

Cho, In-Koo and Thomas J. Sargent, Self-confirming Equilibria, Palgrave Macmillan UK, December 2016.

de la Barrera, Marc and Tim de Silva, “Model-Agnostic Dynamic Programming,” Working

Paper, Stanford University 2024.

DeepSeek-AI, “DeepSeek-R1: Incentivizing Reasoning Capability in LLMs via Reinforcement

Learning,” 2025.

Den Haan, Wouter J., “Heterogeneity, Aggregate Uncertainty, and the Short-Term Interest

Rate,” Journal of Business \& Economic Statistics, October 1996, 14 (4), 399–411.

, Kenneth L. Judd, and Michel Juillard, “Computational suite of models with heterogeneous agents: Incomplete markets and aggregate uncertainty,” Journal of Economic Dynamics

and Control, 2010, 34 (1), 1–3.

Dou, Winston Wei, Itay Goldstein, and Yan Ji, “AI-powered trading, algorithmic collusion,

and price efficiency,” 2025.

Duarte, Victor, Diogo Duarte, and Dejanir H Silva, “Machine Learning for Continuous-Time

Finance,” The Review of Financial Studies, 2024, 37 (11), 3217–3271.

Erev, Ido and Alvin E. Roth, “Learning in Extensive-Form Games: Experimental Data and

Simple Dynamic Models in the Intermediate Term,” Games and Economic Behavior, 1995, 8

(1), 164–212.

and , “Predicting How People Play Games: Reinforcement Learning in Experimental

Games with Unique, Mixed Strategy Equilibria,” American Economic Review, 1998, pp. 848–

881.

Evans, George W. and Seppo Honkapohja, Learning and Expectations in Macroeconomics,

Princeton University Press, 2001.

Farmer, Leland E. and Alexis Akira Toda, “Discretizing nonlinear, non-Gaussian Markov processes with exact conditional moments,” Quantitative Economics, 2017, 8 (2), 651–683.

Favilukis, Jack, Sydney C. Ludvigson, and Stijn Van Nieuwerburgh, “The Macroeconomic

Effects of Housing Wealth, Housing Finance, and Limited Risk Sharing in General Equilibrium,” Journal of Political Economy, 2017, 125 (1), 140–223.

Fernández-Villaverde, Jesús, Galo Nuño, and Jesse Perla, “Taming the Curse of Dimensionality: Quantitative Economics with Deep Learning,” NBER Working Papers 33117, National

Bureau of Economic Research November 2024.

, , and Samuel Hurtado, “Financial Frictions and the Wealth Distribution,” Econometrica,

May 2023, 91 (3), 869–901.

, , Joël Marbet, and Omar Rachedi, “Inequality and the Zero Lower Bound,” Journal of

Econometrics, 2024, p. 105819.

Freeman, C. Daniel, Erik Frey, Anton Raichuk, Sertan Girgin, Igor Mordatch, and Olivier

Bachem, “Brax – A Differentiable Physics Engine for Large Scale Rigid Body Simulation,”

2021.

Fudenberg, Drew and David K. Levine, “Whither Game Theory? Towards a Theory of Learning in Games,” Journal of Economic Perspectives, 2016, 30 (4), 151–170.

Gabriele, Federico, Aldo Glielmo, and Marco Taboga, “Heterogeneous RBCs via deep multiagent reinforcement learning,” 2025.

Gershman, Samuel J. and Nathaniel D. Daw, “Reinforcement Learning and Episodic Memory

in Humans and Animals: An Integrative Framework,” Annual Review of Psychology, 2017, 68

(Volume 68, 2017), 101–128.

Giusto, Andrea, “Adaptive Learning and Distributional Dynamics in an Incomplete Markets

Model,” Journal of Economic Dynamics and Control, 2014, 40, 317–333.

Glimcher, Paul W., “Understanding Dopamine and Reinforcement Learning: The Dopamine

Reward Prediction Error Hypothesis,” Proceedings of the National Academy of Sciences, 2011,

108 (supplement 3), 15647–15654.

, Colin Camerer, Ernst Fehr, and Russell Poldrack, Neuroeconomics: Decision Making and the

Brain Academic Press, Academic Press, 2013.

Gomes, Francisco and Alexander Michaelides, “Asset Pricing with Limited Risk Sharing and

Heterogeneous Agents,” Review of Financial Studies, January 2008, 21 (1), 415–448.

Gopalakrishna, Goutham, Zhouzhou Gu, and Jonathan Payne, “Asset Pricing, Participation

Constraints, and Inequality,” Working Paper 2024.

Gu, Zhouzhou, Mathieu Laurière, Sebastian Merkel, and Jonathan Payne, “Global Solutions

to Master Equations for Continuous Time Heterogeneous Agent Macroeconomic Models,”

arXiv preprint arXiv:2406.13726, 2024.

Guarda, Sebastian, “Narrow and Short Beliefs in Macroeconomics with Heterogeneous

Agents,” Technical Report, Princeton University November 2025. Working paper, available

at https://files.sebastianguarda.com/JMP.pdf.

Han, Jiequn, Yucheng Yang, and Weinan E, “DeepHAM: A Global Solution Method For Heterogeneous Agent Models With Aggregate Shocks,” arXiv preprint arXiv:2112.14377, 2021.

Hausknecht, Matthew and Peter Stone, “Deep Recurrent Q-Learning for Partially Observable

MDPs,” 2017.

Heathcote, Jonathan, Kjetil Storesletten, and Giovanni L. Violante, “Quantitative Macroeconomics with Heterogeneous Households,” Annual Review of Economics, 05 2009, 1 (1), 319–

354.

Hu, Yuanming, Luke Anderson, Tzu-Mao Li, Qi Sun, Nathan Carr, Jonathan Ragan-Kelley,

and Frédo Durand, “DiffTaichi: Differentiable Programming for Physical Simulation,” 2020.

Huang, Ji, “Breaking the Curse of Dimensionality in Heterogeneous-Agent Models: A Deep

Learning-Based Probabilistic Approach,” 2023.

Huggett, Mark, “The risk-free rate in heterogeneous-agent incomplete-insurance economies,”

Journal of Economic Dynamics and Control, 1993, 17 (5-6), 953–969.

Jacobson, Margaret M., “Beliefs, Aggregate Risk, and the U.S. Housing Boom,” Technical Report 2025.

Kahou, Mahdi Ebrahimi, Jesús Fernández-Villaverde, Jesse Perla, and Arnav Sood, “Exploiting symmetry in high-dimensional dynamic programming,” Technical Report 2021.

Kaplan, Greg, Kurt Mitman, and Giovanni L. Violante, “The Housing Boom and Bust: Model

Meets Evidence,” Journal of Political Economy, 2020, 128 (9), 3285–3345.

Kase, Hanno, Leonardo Melosi, and Matthias Rottner, “Estimating Nonlinear Heterogeneous

Agent Models with Neural Networks,” Working Paper, FRB Chicago 2024.

Krueger, D., K. Mitman, and F. Perri, “Chapter 11 - Macroeconomics and Household Heterogeneity,” in John B. Taylor and Harald Uhlig, eds., John B. Taylor and Harald Uhlig, eds., Vol. 2

of Handbook of Macroeconomics, Elsevier, 2016, pp. 843–921.

Krusell, Per and Anthony A. Smith, “Income and wealth heterogeneity, portfolio choice, and

equilibrium asset returns,” Macroeconomic dynamics, 1997, 1 (2), 387–422.

and , “Income and Wealth Heterogeneity in the Macroeconomy,” Journal of Political Economy, October 1998, 106 (5), 867–896.

and , Quantitative Macroeconomic Models with Heterogeneous Agents Econometric Society

Monographs, Cambridge University Press, 2006.

Lasry, Jean-Michel and Pierre-Louis Lions, “Mean field games,” Japanese Journal of Mathematics, 2007, 2, 229–260.

Laurière, Mathieu, Sarah Perrin, Julien Pérolat, Sertan Girgin, Paul Muller, Romuald Élie,

Matthieu Geist, and Olivier Pietquin, “Learning in Mean Field Games: A Survey,” 2024.

, , Sertan Girgin, Paul Muller, Ayush Jain, Theophile Cabannes, Georgios Piliouras,

Julien Perolat, Romuald Elie, Olivier Pietquin, and Matthieu Geist, “Scalable Deep Reinforcement Learning Algorithms for Mean Field Games,” in Kamalika Chaudhuri, Stefanie

Jegelka, Le Song, Csaba Szepesvari, Gang Niu, and Sivan Sabato, eds., Proceedings of the

39th International Conference on Machine Learning, Vol. 162 of Proceedings of Machine Learning

Research PMLR 17–23 Jul 2022, pp. 12078–12095.

Lee, Donghoon and Kenneth I. Wolpin, “Intersectoral Labor Mobility and the Growth of the

Service Sector,” Econometrica, 2006, 74 (1), 1–46.

Llull, Joan, “Immigration, Wages, and Education: A Labour Market Equilibrium Structural

Model,” The Review of Economic Studies, 2018, 85 (3), 1852–1896.

Maliar, Lilia and Serguei Maliar, “Deep learning: Solving HANC and HANK models in the

absence of Krusell-Smith aggregation,” Available at SSRN 3758315, 2020.

, , and Pablo Winant, “Deep Learning for Solving Dynamic Economic Models,” Journal of

Monetary Economics, 2021, 122 (C), 76–101.

Malmendier, Ulrike and Stefan Nagel, “Depression Babies: Do Macroeconomic Experiences

Affect Risk Taking?,” The Quarterly Journal of Economics, 2011, 126 (1), 373–416.

and , “Learning from Inflation Experiences,” The Quarterly Journal of Economics, October

2015, 131 (1), 53–87.

Marcet, Albert and Thomas J. Sargent, “Convergence of Least Squares Learning Mechanisms

in Self-referential Linear Stochastic Models,” Journal of Economic Theory, August 1989, 48 (2),

337–368.

Mnih, Volodymyr, Koray Kavukcuoglu, David Silver, Andrei A. Rusu, Joel Veness, Marc G.

Bellemare, Alex Graves, Martin Riedmiller, Andreas K. Fidjeland, Georg Ostrovski, Stig

Petersen, Charles Beattie, Amir Sadik, Ioannis Antonoglou, Helen King, Dharshan Kumaran, Daan Wierstra, Shane Legg, and Demis Hassabis, “Human-level control through

deep reinforcement learning,” Nature, Feb 2015, 518 (7540), 529–533.

Moll, Benjamin, “The Trouble with Rational Expectations in Heterogeneous Agent Models: A

Challenge for Macroeconomics,” 2025.

Murphy, Kevin, “Reinforcement Learning: An Overview,” 2025.

Niv, Yael, “Reinforcement Learning in the Brain,” Journal of Mathematical Psychology, 2009, 53

(3), 139–154. Special Issue: Dynamic Decision Making.

OpenAI, “OpenAI o1 System Card,” 2024.

Payne, Jonathan, Adam Rebei, and Yucheng Yang, “Deep Learning for Search and Matching

Models,” 2024.

Quadrini, Vincenzo and José-Vı́ctor Rı́os-Rull, “Inequality in Macroeconomics,” Handbook of

Income Distribution, 2015, 2, 1229 – 1302.

Rafailov, Rafael, Archit Sharma, Eric Mitchell, Stefano Ermon, Christopher D. Manning,

and Chelsea Finn, “Direct Preference Optimization: Your Language Model is Secretly a

Reward Model,” 2024.

Raissi, M., P. Perdikaris, and G.E. Karniadakis, “Physics-informed neural networks: A deep

learning framework for solving forward and inverse problems involving nonlinear partial

differential equations,” Journal of Computational Physics, 2019, 378, 686–707.

Robbins, Herbert and Sutton Monro, “A Stochastic Approximation Method,” The Annals of

Mathematical Statistics, 1951, 22 (3), 400–407.

Rotemberg, Julio J, “Sticky prices in the United States,” Journal of political economy, 1982, 90 (6),

1187–1211.

Sargent, Thomas J., “Equilibrium with Signal Extraction from Endogenous Variables,” Journal

of Economic Dynamics and Control, 1991, 15 (2), 245–273.

, The Conquest of American Inflation, Princeton University Press, 1999.

, “HAOK and HANK Models,” Unpublished manuscript, 2023. Available at http://www.

tomsargent.com/research/HAOK\_HANK.pdf.

and John Stachurski, “Quantitative Economics with JAX,” 2025. Available at https://jax.

quantecon.org/intro.html.

Schaab, Andreas, “Micro and Macro Uncertainty,” Working Paper, UC Berkeley 2020.

Silver,

David,

Introduction to Reinforcement Learning,

DeepMind x UCL,

Video

recordings

available

at

https://www.youtube.com/playlist?list=

PLqYmG7hTraZDM-OYHWgPebj2MfCFzFObQ, 2015.

, Aja Huang, Chris J. Maddison, Arthur Guez, Laurent Sifre, George van den Driessche,

Julian Schrittwieser, Ioannis Antonoglou, Veda Panneershelvam, Marc Lanctot, Sander

Dieleman, Dominik Grewe, John Nham, Nal Kalchbrenner, Ilya Sutskever, Timothy Lillicrap, Madeleine Leach, Koray Kavukcuoglu, Thore Graepel, and Demis Hassabis, “Mastering the game of Go with deep neural networks and tree search,” Nature, Jan 2016, 529

(7587), 484–489.

, Julian Schrittwieser, Karen Simonyan, Ioannis Antonoglou, Aja Huang, Arthur Guez,

Thomas Hubert, Lucas Baker, Matthew Lai, Adrian Bolton, Yutian Chen, Timothy Lillicrap, Fan Hui, Laurent Sifre, George van den Driessche, Thore Graepel, and Demis Hassabis, “Mastering the Game of Go without Human Knowledge,” Nature, Oct 2017, 550 (7676),

354–359.

Storesletten, Kjetil, Chris Telmer, and Amir Yaron, “Asset Pricing with Idiosyncratic Risk

and Overlapping Generations,” Review of Economic Dynamics, October 2007, 10 (4), 519–548.

Sutton, Richard S. and Andrew G. Barto, Reinforcement Learning: An Introduction, MIT Press,

2018.

Wu, Zida, Mathieu Lauriere, Matthieu Geist, Olivier Pietquin, and Ankur Mehta,

“Population-aware Online Mirror Descent for Mean-Field Games with Common Noise by

Deep Reinforcement Learning,” 2025.

Xu, Ruitu, Yifei Min, Tianhao Wang, Michael I. Jordan, Zhaoran Wang, and Zhuoran

Yang, “Finding Regularized Competitive Equilibria of Heterogeneous Agent Macroeconomic Models via Reinforcement Learning,” in Francisco Ruiz, Jennifer Dy, and Jan-Willem

van de Meent, eds., Proceedings of The 26th International Conference on Artificial Intelligence and

Statistics, Vol. 206 of Proceedings of Machine Learning Research PMLR 25–27 Apr 2023, pp. 375–

407.

Yang, Yaodong, Rui Luo, Minne Li, Ming Zhou, Weinan Zhang, and Jun Wang, “Mean Field

Multi-Agent Reinforcement Learning,” in Jennifer Dy and Andreas Krause, eds., Proceedings of the 35th International Conference on Machine Learning, Vol. 80 of Proceedings of Machine

Learning Research PMLR 10–15 Jul 2018, pp. 5571–5580.

Young, Eric R., “Solving the incomplete markets model with aggregate uncertainty using the

Krusell–Smith algorithm and non-stochastic simulations,” Journal of Economic Dynamics and

Control, 2010, 34 (1), 36–41.

Zhao,

Shiyu,

Mathematical

Foundations

of

Reinforcement

Learning,

Springer

Nature,

available

at

https://github.com/MathFoundationRL/

Book-Mathematical-Foundation-of-Reinforcement-Learning, 2025.

